---
title: "Control, A, B, and A+B Designs with Withdrawal and Carryover"
author: "Amanda Li"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Control, A, B, and A+B Designs with Withdrawal and Carryover}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette illustrates the component engine introduced in version 0.4.0.
The schedule contains four assigned treatment states: Control, A, B, and A+B.
Unlike a cumulative rollout, a component may be withdrawn and later restarted.
The data-generating model can retain a prespecified residual component effect
after withdrawal.

## Schedule

The following example contains both `Control -> A -> A+B -> B` and
`Control -> B -> A+B` paths. Numeric state codes are 0 for Control, 1 for A, 2
for B, and 3 for A+B.

```{r design}
library(stepwedgepower)

state <- rbind(
  S1 = c(0, 0, 1, 1, 3, 3, 2, 2, 2, 2),
  S2 = c(0, 0, 0, 1, 1, 3, 3, 2, 2, 2),
  S3 = c(0, 0, 2, 2, 3, 3, 3, 3, 3, 3),
  S4 = c(0, 0, 0, 2, 2, 3, 3, 3, 3, 3),
  S5 = c(0, 0, 1, 1, 1, 3, 3, 3, 3, 3),
  S6 = c(0, 0, 2, 2, 2, 3, 3, 3, 3, 3),
  S7 = c(0, 0, 0, 0, 0, 1, 1, 3, 2, 2),
  S8 = c(0, 0, 0, 0, 0, 2, 2, 3, 1, 1)
)

design <- sw_component_design(
  clusters_per_sequence = rep(5L, 8),
  state = state
)
design
```

The schedule is represented internally by separate binary assignment matrices
for A and B. It is not treated as an ordinal 0--3 dose.

## Effects, wash-in, and carryover

```{r assumptions}
assumptions <- sw_component_assumptions(
  baseline_prob = 0.15,
  treatment_or_a = 1.35,
  treatment_or_b = 1.25,
  interaction_or = 1.10,
  delay_a = 1L,
  delay_b = 1L,
  carryover_periods_a = 2L,
  carryover_periods_b = 2L,
  carryover_weights_a = c(0.60, 0.30),
  carryover_weights_b = c(0.50, 0.25),
  restart_rule_a = "reset",
  restart_rule_b = "reset",
  interaction_mode = "effective_overlap",
  icc = 0.05,
  period_effects = log(seq(1.00, 1.10, length.out = ncol(state))),
  n_per_cluster_period = 25L
)
assumptions
```

A one-period delay means that a component first contributes its full effect in
the second consecutive assigned period. After A is withdrawn, 60 percent and
30 percent of its full log-odds effect persist for two periods. With
`interaction_mode = "effective_overlap"`, the interaction weight is the product
of the effective A and B weights, so interaction carryover is possible while
both effective components overlap.

## Design audit

```{r audit}
audit <- audit_component_design(design, assumptions)
audit
```

The audit checks the rank of the intended fixed-effect model and formal
estimability of six standard contrasts. It also reports concurrent state
comparisons and notes any withdrawal for which immediate washout has been
assumed.

## One simulated trial

```{r simulate-fit}
one_trial <- simulate_component_swcrt(design, assumptions, seed = 123)
fit <- fit_component_model(
  one_trial,
  multiplicity = "holm",
  multiplicity_family = c("A_vs_control", "B_vs_control", "interaction")
)
fit$contrasts
```

The standard contrasts are:

* A versus Control;
* B versus Control;
* A+B versus Control;
* A+B versus A;
* A+B versus B; and
* the A-by-B interaction.

The interaction coefficient is the departure from additivity on the log-odds
scale. A B-only state is required to separate the B main effect from the
interaction when the interaction is fitted.

## Power and global-null calibration

A short exploratory run is shown below. A final analysis should use a simulation
count chosen from a prespecified Monte Carlo precision target and should also
run the global-null design through `type1_component_swcrt()`.

```{r power, eval=FALSE}
power <- power_component_swcrt(
  design,
  assumptions,
  nsim = 5000,
  multiplicity = "holm",
  multiplicity_family = c("A_vs_control", "B_vs_control", "interaction"),
  nAGQ = 1,
  n_cores = 4,
  seed = 2026
)
summary(power)

type1 <- type1_component_swcrt(
  design,
  assumptions,
  nsim = 5000,
  multiplicity = "holm",
  multiplicity_family = c("A_vs_control", "B_vs_control", "interaction"),
  nAGQ = 1,
  n_cores = 4,
  seed = 22026
)
summary(type1)
```

Both conditional power and failure-aware power are reported. Their difference
shows how much the apparent operating performance depends on excluding failed
or non-estimable analyses.
