---
title: "Structurally Incomplete and Block Stepped-Wedge Designs"
author: "Amanda Li"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structurally Incomplete and Block Stepped-Wedge Designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette compares a conventional stepped-wedge design (SWD) with a block
stepped-wedge design (BSWD). A dash denotes a clinic-period for which no outcome
is observed. It is not a Control observation.

## Exact schedules

The manuscript labels are `0` for Control, `1` for intervention A, and `1+2`
for A+B. The component engine maps `1+2` to its internal A+B state.

```{r designs}
library(stepwedgepower)

swd <- sw_incomplete_component_design(
  clusters_per_sequence = c(5L, 5L),
  state = rbind(
    `Group 1` = c("0", "1", "1+2", "1+2"),
    `Group 2` = c("0", "0", "1", "1+2")
  )
)

bswd <- sw_incomplete_component_design(
  clusters_per_sequence = c(2L, 2L, 3L, 3L),
  state = rbind(
    `Group 1` = c("0", "1", "1+2", "1+2", "-"),
    `Group 2` = c("0", "0", "1", "1+2", "-"),
    `Group 3` = c("-", "0", "1", "1+2", "1+2"),
    `Group 4` = c("-", "0", "0", "1", "1+2")
  )
)

swd
bswd
```

For leading missing periods, the constructor infers the first observed state as
the latent state. For trailing missing periods, it carries forward the final
observed state. These latent values preserve treatment history but do not
create outcome observations. An internal missing gap requires the user to
supply `latent_state` explicitly.

## Assumptions and equal observed resources

B is not observed by itself in these designs. With no interaction term, the B
coefficient is therefore the incremental A+B versus A effect.

```{r assumptions-resources}
assumptions <- sw_component_assumptions(
  baseline_prob = 0.15,
  treatment_or_a = 1.35,
  treatment_or_b = 1.25,
  interaction_mode = "none",
  icc = 0.05,
  n_per_cluster_period = 25L
)

rbind(
  SWD = component_resource_summary(swd, assumptions),
  BSWD = component_resource_summary(bswd, assumptions)
)
```

The SWD contains 40 calendar and observed clinic-periods. The BSWD contains 50
calendar clinic-periods, of which 40 are observed and 10 are structurally
missing. Both designs therefore contain 1,000 individual observations when
there are 25 observations per clinic-period.

## Audit and simulation

```{r audit-simulate}
contrasts <- c("A_vs_control", "AB_vs_A", "AB_vs_control")

audit_component_design(
  bswd,
  assumptions,
  contrasts = contrasts,
  include_interaction = FALSE
)

observed_trial <- simulate_component_swcrt(bswd, assumptions, seed = 123)
latent_trial <- simulate_component_swcrt(
  bswd, assumptions, seed = 123, include_unobserved = TRUE
)

c(
  observed_rows = nrow(observed_trial),
  latent_calendar_rows = nrow(latent_trial),
  structurally_unobserved_rows = sum(!latent_trial$observed)
)
```

By default, the simulator returns only observed cluster-periods. Setting
`include_unobserved = TRUE` retains the latent rows for inspection; their
sample size and event count are missing, and `fit_component_model()` excludes
them automatically.

## Power comparison

The same component power engine can compare complete and structurally
incomplete designs. The comparison reports both calendar and observed
resources.

```{r power, eval=FALSE}
comparison <- compare_component_designs(
  designs = list(SWD = swd, BSWD = bswd),
  assumptions = assumptions,
  nsim = 5000,
  contrasts = contrasts,
  include_interaction = FALSE,
  multiplicity = "holm",
  multiplicity_family = contrasts,
  nAGQ = 1,
  n_cores = 4,
  seed = 2026,
  warn_on_design = FALSE
)

comparison$resource_check
comparison$comparison
```

The complete installed demonstration can be located with:

```{r locate-demo}
system.file("examples", "demo_exact_SWD_BSWD.R", package = "stepwedgepower")
```
